Supplementary Material

### 2020-04-24 Revisions2/Beauchesne-eDriversMS-SI.rmd
finish answer to reviewer comments

<!--
rmarkdown::render('./Revisions/Beauchesne-eDriversMS-SI.rmd')
-->

@@

Table S1. List of drivers currently available on *eDrivers* along with their respective acronym used in the figures \textcolor{red}{\sout{in}} \textcolor{blue}{of} the supplementary material.

@@

Oceans are the largest reservoir\textcolor{blue}{s} and sink\textcolor{blue}{s} of atmospheric carbon dioxide ($CO_2$). Its uptake increases seawater acidity and \textcolor{red}{\sout{the}} lowers \textcolor{blue}{the} saturation state \textcolor{blue}{of waters} with respect to calcite ($\Omega_C$) and aragonite ($\Omega_A$), the two most common $CaCO_3$ polymorphs \textcolor{blue}{that constitute} \textcolor{red}{\sout{constituting}} the shells and skeleton\textcolor{blue}{s} of many marine organisms [@mucci2017]. \textcolor{blue}{Ocean acidification} \textcolor{red}{\sout{This}} can have deleterious effects on carbonate-secreting organisms (*e.g.* mollusks and crustaceans) and certain physiological processes in non-calcifying organisms [@fabry2008; @kroeker2013].

When $\Omega_C$ or $\Omega_A$ decrease below 1, water becomes undersaturated and corrosive to the skeletal minerals of carbonate-secreting organisms. Note that if $\Omega_A$ $< 1$, the waters may still be supersaturated with respect to calcite since it is 50% more soluble than aragonite. However, organisms have wide ranging responses to \textcolor{blue}{changes in $\Omega_C$ and $\Omega_A$} \textcolor{red}{\sout{saturation state}}. For example, most corals stop calcifying at $\Omega_A$ $< 2$ [@kleypas2006; @bove2019], while other organisms may have adapted to precipitate $CaCO_3$ even when seawater is undersaturated with respect to calcite [$\Omega_C$ $< 1$ or $\Omega_A$ $< 0.65$; e.g. @uthicke2016].

We used $\Omega_{A}$ \textcolor{red}{\sout{saturation state}} to characterize ocean acidification in the bottom waters of the St. Lawrence. The data come from the Department of Fisheries and Oceans' (DFO) Atlantic Zone Monitoring Program [AZMP; @galbraith2018] surveys and DFO's multispecies surveys [@bourdages2018] collected in August-September of 2017 [@starr2019]. \textcolor{blue}{$\Omega_{A}$, pH, and dissolved oxygen ($O_2$) were measured at 117 stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The majority of sampling stations were located in the Southern Gulf ($n = 84$), the Estuary and Northern Gulf ($n = 33$).}

The carbonate \textcolor{red}{\sout{chemistry}} \textcolor{blue}{speciation} was determined through pH and \textcolor{blue}{total alkalinity} (TA) measurements. Samples for pH and \textcolor{red}{\sout{total alkalinity}} TA were collected \textcolor{red}{\sout{under a bubble free and}} \textcolor{blue}{with} no head space \textcolor{red}{\sout{conditions}} in\textcolor{red}{\sout{to}} 500 $mL$ borosilicate glass flasks\textcolor{red}{\sout{.}}\textcolor{blue}{, and} 250 $\mu L$ of \textcolor{red}{\sout{the}} saturated $HgCl_2$ solution was
added to \textcolor{blue}{each} sample\textcolor{red}{\sout{,}} and processed following the "Guide to best practices for Ocean
CO2 Measurements" [@dickson2007].

$pH_T$ was determined spectrophotometrically using the indicator dye m-cresol
purple (Sigma-Aldrich). Absorbance was measured at 730, 578 and 434 nm before
and after dye addition in 10 cm quartz cells thermostated at 25 ± $0.05^oC$
[@dickson2007]. A similar procedure was carried out before each set of sample
measurement\textcolor{red}{\sout{s}} using a TRIS (Tris (hydroxymethyl) -aminomethane) buffer prepared
at a practical salinity (S) of approximately 30 [@millero1986]. Certified
Reference Material (CRM) (supplied by Professor Andrew Dickson, Scripps
Institution of Oceanography, San Diego, USA) was used for quality control of
our pH TRIS buffer.

TA was determined by potentiometric titration in an open cell using an
automated Radiometer potentiometric titrator (Titrilab 865) and a pH combination
electrode (pHC2001) in a continuous titrant addition mode, an algorithm
specifically designed for shallow end-point detection [@dickson2007]. The
\textcolor{red}{\sout{dilute}} \textcolor{blue}{0.1M} HCl titrant
\textcolor{red}{\sout{0.1M}} in \textcolor{red}{\sout{a solution of}}
\textcolor{blue}{0.6M} NaCl \textcolor{red}{\sout{of 0.6M}} was calibrated using CRM provided by Professor Andrew G. Dickson.

The carbonate system parameters (including $\Omega_{A}$) were
corrected for *in situ* pressure and temperature using the algorithm CO2SYS
[@lewis1998] with measured pH, \textcolor{red}{\sout{total alkalinity}} \textcolor{blue}{TA}, soluble reactive \textcolor{blue}{phosphate}
and silicate concentrations as input parameters.
\textcolor{red}{\sout{$\Omega_{A}$, pH, dissolved oxygen ($O_2$) and pressure were sampled at 117
stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The
majority of sampling stations were located in the Southern Gulf ($n = 84$),
compared to the Estuary and Northern Gulf ($n = 33$).}}

To account for \textcolor{blue}{the} low sample size in the Northern Gulf, we used the correlation
between $O_2$ and $CO_2$, which are linked through the stoichiometry of the
respiration reaction in waters that are isolated from the atmosphere [@mucci2011].
Metabolic $CO_2$ increases in water through biotic processes such microbial
respiration of organic matter [@mucci2011]. In deep waters\textcolor{blue}{, found in the Laurentian, Anticosti and Esquiman Channels}, variations of
$\Omega_{A}$ are largely governed by the addition of metabolic $CO_2$, whereas\textcolor{blue}{,} near the air-sea interface\textcolor{blue}{,} $O_2$ is replenished faster than $CO_2$ can escape
[@zeebe2001]. The impact of respiration on carbonate chemistry is highest in
hypoxic regions, where metabolic $CO_2$ accumulates to high concentrations
[@mucci2011].

We explored the correlation between $\Omega_{A}$ and $O_2$ in the data and found
no significant correlation at the scale of the whole St. Lawrence (*p-value*:
0.08; $R^2$: 0.03). Nevertheless, a correlation between $\Omega_A$ and $O_2$
exists in the Northern St. Lawrence (*p-value*: < 0.01; $R^2$: 0.49), but not
in the Southern St. Lawrence (*p-value*: 0.85; $R^2$: < 0.01). Furthermore, the
correlation observed in the north between $\Omega_A$ and $O_2$ is especially
strong when deep Channel\textcolor{red}{\sout{s}} \textcolor{blue}{waters} (*Deep* model: *p-value*: < 0.01; $O_2$: 0.20; $R^2$:
0.97) are separated from shallower regions (*Shallow* model: *p-value*: < 0.01;
$O_2$: 0.07; $R^2$: 0.96), with $\Omega_A$ increasing more steeply with
increasing $O_2$ in \textcolor{blue}{the} deep Channels. This stronger correlation is likely caused by the lack of gas exchange with the atmosphere or the surface mixed layer and the importance of metabolic $CO_2$ in deep waters, some of which may have mean ages of up to 20 years [@mucci2011]. In the Northern Gulf, the acidification \textcolor{red}{\sout{is following}} \textcolor{blue}{trends with} depth and distance from the Cabot Strait along the Laurentian, Anticosti and Esquiman Channels due to the progressive oxygen depletion and metabolic carbon dioxide accumulation [@mucci2011].

There are likely different dynamics at play in the Southern Gulf, such as waters with lowered alkalinity and salinity caused by the \textcolor{blue}{surface, seaward-flowing Gaspé Current} \textcolor{red}{\sout{St. Lawrence plume}}. \textcolor{red}{\sout{Such}} \textcolor{blue}{These} water\textcolor{blue}{s} \textcolor{red}{\sout{masses}} \textcolor{red}{\sout{are}} have lower buffering capacities to decreasing pH and $\Omega_{A}$. \textcolor{red}{\sout{Also}} \textcolor{blue}{Furthermore}, the accumulation of metabolic carbon dioxide produced by rapid oxic degradation of organic matter in the sandy and permeable sediments of the Magdalen Shelf or \textcolor{red}{\sout{through}} \textcolor{blue}{in} the \textcolor{red}{\sout{St. Lawrence plume}} \textcolor{blue}{Gaspé current} could decrease $\Omega_{A}$ in the Southern Gulf (Starr and Chassé, *personal communication*).

We thus divided the \textcolor{red}{\sout{analyses}} \textcolor{blue}{and their analysis} between the Southern and Northern St. Lawrence. The St. Lawrence was divided \textcolor{red}{\sout{using}} \textcolor{blue}{along} the 350m isobath on the southern slope of the Laurentian Channel, from the Cabot Strait to the tip of the Gaspé Peninsula. \textcolor{blue}{The} $\Omega_A$ distribution in the Southern St. Lawrence was modeled
using the exponential kriging model.

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. Whereas
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentration\textcolor{blue}{s} in the water column \textcolor{red}{\sout{from}} \textcolor{blue}{available through}
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2019] and interpolated oxygen % saturation using cokriging with depth as a covariable, as done in @dutil2011. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia \textcolor{blue}{data layer}
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

Based on the wide range of $\Omega_A$ levels at which organism\textcolor{blue}{s} are negatively
affected (see above), we built an index of acidification stress ($A_s$)
ranging between 0 and 1 that slowly \textcolor{red}{\sout{begins increasing}} \textcolor{blue}{increases} as $\Omega_A$ \textcolor{blue}{decreases to} $\approx$
$2$, increases more rapidly close to $\Omega_A$ $= 1$ and reaches its peak at
$\Omega_A$ $\approx$ $0.5$:

@@

The data used to characterize hypoxia
\textcolor{red}{\sout{uses oxygen saturation in \% and}}
come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] survey\textcolor{red}{\sout{s}} in late spring and fall
\textcolor{blue}{of 2013 to 2017}, as well as from \textcolor{red}{\sout{annual}}
DFO’s \textcolor{blue}{annual} multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. We provide a brief summary of data and methods to
describe hypoxia in this document. For more details, refer to @blais2019.

Oxygen concentration \textcolor{red}{\sout{is}} \textcolor{blue}{were} measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on \textcolor{red}{\sout{the}} \textcolor{blue}{a} CTD. The probe is calibrated against seawater
samples \textcolor{red}{\sout{that are}} \textcolor{blue}{taken directly from the Niskin bottles at every cast and} analyzed by Winkler titration \textcolor{red}{\sout{on every cast}} [see @blais2019 for calibration procedure]. The data used is the last depth
sampled on the CTD profile, which \textcolor{blue}{is} typically \textcolor{red}{\sout{corresponds to}} ~10m above the
bottom. \textcolor{blue}{Note that this depth does not necessarily correspond to the depth of the oxygen minimum. The oxygen minimum is typically found along the 27.25 isopycnal, often situated as much as 100 m above bottom in the Lower Estuary.} Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation\textcolor{blue}{,} taking into account the salinity and temperature at the selected
depth.

We interpolated oxygen saturation using cokriging with depth as a covariable,
as done in @dutil2011. Bathymetric data come from @dutil2011 and have a
100 $km^2$ resolution. According to @diaz1995, severe hypoxia \textcolor{red}{\sout{can be observed}} \textcolor{blue}{occurs}
when dissolved oxygen falls below 2 $ml$ $L^{-1}$, \textcolor{red}{\sout{which corresponds}} to 62.5 $\mu mol$ $L^{-1}$ \textcolor{red}{\sout{and}} \textcolor{blue}{or} ~20% saturation. This is considered as the level
necessary to maintain most animal life. \textcolor{red}{\sout{Similarly}} \textcolor{blue}{Accordingly}, @chabot2008 studied the
effects of hypoxia on the energy budget of cod in the St. Lawrence and found
that behavioural effects began manifesting \textcolor{blue}{themselves} below 70% oxygen saturation and that
survival becomes jeopardized below ~20% saturation. We used these observations
to create an index of hypoxia using an inverted logistic curve that slowly
increases below 70% oxygen saturation\textcolor{red}{\sout{,}} \textcolor{blue}{and} increases acutely as it \textcolor{red}{\sout{approaches the 20\% oxygen saturation threshold and}} reaches its peak intensity between 20 and 30% oxygen saturation:

@@

### Bottom\textcolor{blue}{-water} temperature anomalies

The data used to characterize bottom\textcolor{blue}{-water} temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods to
characterize the bottom\textcolor{blue}{-water} temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

Bottom\textcolor{blue}{-water} temperatures are interpolated in the Gulf using
conductivity-temperature-depth (CTD) sampling performed annually through DFO's
multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. Using this sampling survey, temperatures are horizontally
interpolated at each 1 m depth layer on a 2 km resolution grid. Bottom\textcolor{blue}{-water}
temperatures are then extracted by using a bathymetry layer from the Canadian
Hydrographic Survey [@dutil2012] and selecting the interpolated temperature
from the layer corresponding to the bottom depth at each grid point.

We used temperature anomalies, *i.e.* deviations from long-term normal
conditions \textcolor{blue}{between 1981 and 2010}, to measure an annual index of stress associated with extreme
temperatures between 2013 and 2017. Temperature anomalies were calculated using
the difference between grid cell values with 1981-2010 climatological averages.
Anomaly time series were normalized by their standard deviation (SD) to allow
comparison\textcolor{blue}{s} between areas of the St. Lawrence with different temperature ranges.
For example, temperatures observed in deep channels are less variable than in
shallower regions of the St. Lawrence. Hence, if anomalies were expressed in
degrees Celsius, it would underestimate the relative importance of anomalies
in deep channels when compared to shallower regions. Grid cells whose monthly
value exceeded ±0.5 standard deviation (SD) from the long-term average were
considered as anomalous [@galbraith2018]. Outliers in the data were defined as
those that fell beyond the interquartile range * 3, identified as extreme
outliers by @tukey1977. Outlier values were capped to correspond to the 5th and
95th percentile\textcolor{red}{\sout{s}} \textcolor{blue}{values}. Anomalies were divided into positive and negative anomalies
and the absolute value of anomalies were used as an indicator of the intensity
of bottom\textcolor{blue}{-water} temperature anomalies. The mean anomaly intensity between 2013 and
2017 for each grid cell was used to generate the final index of bottom\textcolor{blue}{-water}
temperature anomalies.

### Surface\textcolor{blue}{-water} temperature anomalies

The data used to characterize surface\textcolor{blue}{-water} temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods to
characterize the surface\textcolor{blue}{-water} temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

The surface\textcolor{blue}{-water} layer is characterized using sea surface\textcolor{blue}{-water} temperature (SST) monthly
composites from Advanced Very High Resolution Radiometer (AVHRR) satellite
images obtained from the National Oceanic and Atmospheric Administration (NOAA)
and European Organization for the Exploitation of Meteorological Satellites
(EUMETSAT). Images used are from DFO’s Maurice Lamontagne Institute at a 1 km
resolution from 1985-2013 and from DFO’s Bedford Institute of Oceanography (BIO)
Operational Remote Sensing group at a 1.5 km resolution since 2014. Monthly
anomalies were constructed as the difference between monthly averages and the
1985-2010 climatological mean for each month.

Surface\textcolor{blue}{-water} temperature anomalies were characterized following the same method used
for bottom\textcolor{blue}{-water} temperature anomalies. Only the months of May to November were
included to avoid biases associated with the presence of ice cover. Monthly
anomalies from May to November values were summed to obtain an indicator of
annual surface\textcolor{blue}{-water} temperature anomaly intensity in each grid cell. The mean
anomaly intensity between 2013 and 2017 for each grid cell was used to generate
the final index of surface\textcolor{blue}{-water} temperature anomalies.

### Sea\textcolor{blue}{-}level rise

The data used to characterize sea\textcolor{blue}{-}level rise risk come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Sea\textcolor{blue}{-}level rise was characterized by @nicholls2010 using NASA's satellite
altimetry data (Topex/Poseidon, Jason-1&2, GFO, ERS-1&2, and Envisat
missions) and available at
http://www.aviso.altimetry.fr/en/data/products/ocean-indicatorsproducts/mean-sea-level/products-images.html

The rate of sea\textcolor{blue}{-}level rise ($mm/year$) was measured between 1992 and 2012 and
transformed as a net change value ($mm$) by multiplying by the number of years
considered. Only positive values were selected under the assumption that
only positive sea\textcolor{blue}{-}level rise is likely to cause environmental stress.

@@

\textcolor{red}{\sout{Aquaculture data comes from a variety of sources in the St. Lawrence because
aquaculture sites are mostly managed at the provincial level.
We therefore had to gather the data on aquaculture sites from the 5 provinces dividing the
St. Lawrence}}
\textcolor{blue}{As aquaculture sites are managed at the provincial level in the St. Lawrence, data on aquaculture sites come from various sources}
[@mapaq2016; @dfo2016a; @aaf2016; @fa2016; @ffa2016].
Invertebrates aquaculture is especially important in the southern and western
Gulf. Fish and algae aquaculture, on the other hand, remains marginal.
Considering this, we only \textcolor{red}{\sout{considered}} \textcolor{blue}{included} invertebrates aquaculture for the
aquaculture driver layer. However, if fish or algae farming were to become more
important, these driver\textcolor{blue}{s} should be incorporated in future analyses as individual
layers, as impacts vary between types of aquaculture.

Aquaculture activities are highly localized and potential effects do not or
rarely extend beyond the location of the farms. We therefore only considered
the \textcolor{red}{\sout{actual}} location of sites to characterize the distribution of this driver.
We were unable to characterize site production in terms of biomass farmed, which
could provide an indication of the intensity of aquaculture activities. As such,
we considered aquaculture as binary presence-absence data \textcolor{red}{\sout{in our analyses}}.

@@

\textcolor{red}{\sout{We used lights at night as a proxy of coastal infrastructure development, as
terrestrial stable lights at night represent light from human settlements and
industrial sites with electricity.}}
\textcolor{blue}{Terrestrial stable lights at night mostly represent light from human settlements and industrial sites with electricity.
We thus used lights at night as a proxy of coastal infrastructure development.}

@@

As in @halpern2008a and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statistics-canada2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated and they combine
to cover all of Canada. The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

@@

\textcolor{red}{\sout{In order to}} \textcolor{blue}{To} characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multiplied the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

@@

# Threat complexes

>>>>

<!--
rmarkdown::render('./Revisions2/Beauchesne-eDriversMS-SI.rmd')
-->

@@

Table S1. List of drivers currently available on *eDrivers* along with their respective acronym used in the figures of the supplementary material.

@@

Oceans are the largest reservoirs and sinks of atmospheric carbon dioxide ($CO_2$). Its uptake increases seawater acidity and lowers the saturation state of waters with respect to calcite ($\Omega_C$) and aragonite ($\Omega_A$), the two most common $CaCO_3$ polymorphs that constitute the shells and skeletons of many marine organisms [@mucci2017]. Ocean acidification can have deleterious effects on carbonate-secreting organisms (*e.g.* mollusks and crustaceans) and certain physiological processes in non-calcifying organisms [@fabry2008; @kroeker2013].

When $\Omega_C$ or $\Omega_A$ decrease below 1, water becomes undersaturated and corrosive to the skeletal minerals of carbonate-secreting organisms. Note that if $\Omega_A$ $< 1$, the waters may still be supersaturated with respect to calcite since it is 50% more soluble than aragonite. However, organisms have wide ranging responses to changes in $\Omega_C$ and $\Omega_A$. For example, most corals stop calcifying at $\Omega_A$ $< 2$ [@kleypas2006; @bove2019], while other organisms may have adapted to precipitate $CaCO_3$ even when seawater is undersaturated with respect to calcite [$\Omega_C$ $< 1$ or $\Omega_A$ $< 0.65$; e.g. @uthicke2016].

We used $\Omega_{A}$ to characterize ocean acidification in the bottom waters of the St. Lawrence. The data come from the Department of Fisheries and Oceans' (DFO) Atlantic Zone Monitoring Program [AZMP; @galbraith2018] surveys and DFO's multispecies surveys [@bourdages2018] collected in August-September of 2017 [@starr2019]. $\Omega_{A}$, pH, and dissolved oxygen ($O_2$) were measured at 117 stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The majority of sampling stations were located in the Southern Gulf ($n = 84$), the Estuary and Northern Gulf ($n = 33$).

The carbonate speciation was determined through pH and total alkalinity (TA) measurements. Samples for pH and TA were collected with no head space in 500 $mL$ borosilicate glass flasks, and 250 $\mu L$ of saturated $HgCl_2$ solution was added to each sample and processed following the "Guide to best practices for Ocean CO2 Measurements" [@dickson2007].

$pH_T$ was determined spectrophotometrically using the indicator dye m-cresol
purple (Sigma-Aldrich). Absorbance was measured at 730, 578 and 434 nm before
and after dye addition in 10 cm quartz cells thermostated at 25 ± $0.05^oC$
[@dickson2007]. A similar procedure was carried out before each set of sample
measurement using a TRIS (Tris (hydroxymethyl) -aminomethane) buffer prepared
at a practical salinity (S) of approximately 30 [@millero1986]. Certified
Reference Material (CRM) (supplied by Professor Andrew Dickson, Scripps
Institution of Oceanography, San Diego, USA) was used for quality control of
our pH TRIS buffer.

TA was determined by potentiometric titration in an open cell using an
automated Radiometer potentiometric titrator (Titrilab 865) and a pH combination
electrode (pHC2001) in a continuous titrant addition mode, an algorithm
specifically designed for shallow end-point detection [@dickson2007]. The
0.1M HCl titrant in a solution of 0.6M NaCl was calibrated using CRM provided by Professor Andrew G. Dickson.

The carbonate system parameters (including $\Omega_{A}$) were
corrected for *in situ* pressure and temperature using the algorithm CO2SYS
[@lewis1998] with measured pH, TA, soluble reactive phosphate and silicate concentrations as input parameters.

To account for the low sample size in the Northern Gulf, we used the correlation
between $O_2$ and $CO_2$, which are linked through the stoichiometry of the
respiration reaction in waters that are isolated from the atmosphere [@mucci2011].
Metabolic $CO_2$ increases in water through biotic processes such microbial
respiration of organic matter [@mucci2011]. In deep waters, found in the Laurentian, Anticosti and Esquiman Channels, variations of
$\Omega_{A}$ are largely governed by the addition of metabolic $CO_2$, whereas, near the air-sea interface, $O_2$ is replenished faster than $CO_2$ can escape
[@zeebe2001]. The impact of respiration on carbonate chemistry is highest in
hypoxic regions, where metabolic $CO_2$ accumulates to high concentrations
[@mucci2011].

We explored the correlation between $\Omega_{A}$ and $O_2$ in the data and found
no significant correlation at the scale of the whole St. Lawrence (*p-value*:
0.08; $R^2$: 0.03). Nevertheless, a correlation between $\Omega_A$ and $O_2$
exists in the Northern St. Lawrence (*p-value*: < 0.01; $R^2$: 0.49), but not
in the Southern St. Lawrence (*p-value*: 0.85; $R^2$: < 0.01). Furthermore, the
correlation observed in the north between $\Omega_A$ and $O_2$ is especially
strong when deep Channel waters (*Deep* model: *p-value*: < 0.01; $O_2$: 0.20; $R^2$:
0.97) are separated from shallower regions (*Shallow* model: *p-value*: < 0.01;
$O_2$: 0.07; $R^2$: 0.96), with $\Omega_A$ increasing more steeply with
increasing $O_2$ in the deep Channels. This stronger correlation is likely caused by the lack of gas exchange with the atmosphere or the surface mixed layer and the importance of metabolic $CO_2$ in deep waters, some of which may have mean ages of up to 20 years [@mucci2011]. In the Northern Gulf, the acidification trends with depth and distance from the Cabot Strait along the Laurentian, Anticosti and Esquiman Channels due to the progressive oxygen depletion and metabolic carbon dioxide accumulation [@mucci2011].

There are likely different dynamics at play in the Southern Gulf, such as waters with lowered alkalinity and salinity caused by the surface, seaward-flowing Gaspé Current. These waters have lower buffering capacities to decreasing pH and $\Omega_{A}$. Furthermore, the accumulation of metabolic carbon dioxide produced by rapid oxic degradation of organic matter in the sandy and permeable sediments of the Magdalen Shelf or in the Gaspé current could decrease $\Omega_{A}$ in the Southern Gulf (Starr and Chassé, *personal communication*).

We thus divided the and their analysis between the Southern and Northern St. Lawrence. The St. Lawrence was divided along the 350m isobath on the southern slope of the Laurentian Channel, from the Cabot Strait to the tip of the Gaspé Peninsula. The $\Omega_A$ distribution in the Southern St. Lawrence was modeled
using the exponential kriging model.

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. Whereas
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentrations in the water column available through
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2019] and interpolated oxygen % saturation using cokriging with depth as a covariable, as done in @dutil2011. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia data layer
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

Based on the wide range of $\Omega_A$ levels at which organisms are negatively
affected (see above), we built an index of acidification stress ($A_s$)
ranging between 0 and 1 that slowly increases as $\Omega_A$ decreases to $\approx$ $2$, increases more rapidly close to $\Omega_A$ $= 1$ and reaches its peak at
$\Omega_A$ $\approx$ $0.5$:

@@

The data used to characterize hypoxia come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] survey in late spring and fall of 2013 to 2017, as well as from DFO’s annual multispecies surveys for the northern Gulf in August and for the Magdalen Shallows in September. We provide a brief summary of data and methods to describe hypoxia in this document. For more details, refer to @blais2019.

Oxygen concentration were measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on a CTD. The probe is calibrated against seawater
samples taken directly from the Niskin bottles at every cast and analyzed by Winkler titration [see @blais2019 for calibration procedure]. The data used is the last depth sampled on the CTD profile, which is typically ~10m above the
bottom. Note that this depth does not necessarily correspond to the depth of the oxygen minimum. The oxygen minimum is typically found along the 27.25 isopycnal, often situated as much as 100 m above bottom in the Lower Estuary. Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation, taking into account the salinity and temperature at the selected
depth.

We interpolated oxygen saturation using cokriging with depth as a covariable,
as done in @dutil2011. Bathymetric data come from @dutil2011 and have a
100 $km^2$ resolution. According to @diaz1995, severe hypoxia occurs
when dissolved oxygen falls below 2 $ml$ $L^{-1}$, to 62.5 $\mu mol$ $L^{-1}$ or ~20% saturation. This is considered as the level
necessary to maintain most animal life. Accordingly, @chabot2008 studied the
effects of hypoxia on the energy budget of cod in the St. Lawrence and found
that behavioural effects began manifesting themselves below 70% oxygen saturation and that
survival becomes jeopardized below ~20% saturation. We used these observations
to create an index of hypoxia using an inverted logistic curve that slowly
increases below 70% oxygen saturation and increases acutely as it reaches its peak intensity between 20 and 30% oxygen saturation:

@@

### Bottom-water temperature anomalies

The data used to characterize bottom-water temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods to
characterize the bottom-water temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

Bottom-water temperatures are interpolated in the Gulf using
conductivity-temperature-depth (CTD) sampling performed annually through DFO's
multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. Using this sampling survey, temperatures are horizontally
interpolated at each 1 m depth layer on a 2 km resolution grid. Bottom-water
temperatures are then extracted by using a bathymetry layer from the Canadian
Hydrographic Survey [@dutil2012] and selecting the interpolated temperature
from the layer corresponding to the bottom depth at each grid point.

We used temperature anomalies, *i.e.* deviations from long-term normal
conditions between 1981 and 2010, to measure an annual index of stress associated with extreme
temperatures between 2013 and 2017. Temperature anomalies were calculated using
the difference between grid cell values with 1981-2010 climatological averages.
Anomaly time series were normalized by their standard deviation (SD) to allow
comparisons between areas of the St. Lawrence with different temperature ranges.
For example, temperatures observed in deep channels are less variable than in
shallower regions of the St. Lawrence. Hence, if anomalies were expressed in
degrees Celsius, it would underestimate the relative importance of anomalies
in deep channels when compared to shallower regions. Grid cells whose monthly
value exceeded ±0.5 standard deviation (SD) from the long-term average were
considered as anomalous [@galbraith2018]. Outliers in the data were defined as
those that fell beyond the interquartile range * 3, identified as extreme
outliers by @tukey1977. Outlier values were capped to correspond to the 5th and
95th percentile values. Anomalies were divided into positive and negative anomalies and the absolute value of anomalies were used as an indicator of the intensity of bottom-water temperature anomalies. The mean anomaly intensity between 2013 and 2017 for each grid cell was used to generate the final index of bottom-water temperature anomalies.

### Surface-water temperature anomalies

The data used to characterize surface-water temperature anomalies come from the
Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018]. We provide a brief summary of data and methods to
characterize the surface-water temperature climatology and anomalies in this document.
For more details, refer to @galbraith2018.

The surface-water layer is characterized using sea surface-water temperature (SST) monthly
composites from Advanced Very High Resolution Radiometer (AVHRR) satellite
images obtained from the National Oceanic and Atmospheric Administration (NOAA)
and European Organization for the Exploitation of Meteorological Satellites
(EUMETSAT). Images used are from DFO’s Maurice Lamontagne Institute at a 1 km
resolution from 1985-2013 and from DFO’s Bedford Institute of Oceanography (BIO)
Operational Remote Sensing group at a 1.5 km resolution since 2014. Monthly
anomalies were constructed as the difference between monthly averages and the
1985-2010 climatological mean for each month.

Surface-water temperature anomalies were characterized following the same method used
for bottom-water temperature anomalies. Only the months of May to November were
included to avoid biases associated with the presence of ice cover. Monthly
anomalies from May to November values were summed to obtain an indicator of
annual surface-water temperature anomaly intensity in each grid cell. The mean
anomaly intensity between 2013 and 2017 for each grid cell was used to generate
the final index of surface-water temperature anomalies.

### Sea-level rise

The data used to characterize sea-level rise risk come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Sea-level rise was characterized by @nicholls2010 using NASA's satellite
altimetry data (Topex/Poseidon, Jason-1&2, GFO, ERS-1&2, and Envisat
missions) and available at
http://www.aviso.altimetry.fr/en/data/products/ocean-indicatorsproducts/mean-sea-level/products-images.html

The rate of sea-level rise ($mm/year$) was measured between 1992 and 2012 and
transformed as a net change value ($mm$) by multiplying by the number of years
considered. Only positive values were selected under the assumption that
only positive sea-level rise is likely to cause environmental stress.

@@

As aquaculture sites are managed at the provincial level in the St. Lawrence, data on aquaculture sites come from various sources
[@mapaq2016; @dfo2016a; @aaf2016; @fa2016; @ffa2016].
Invertebrates aquaculture is especially important in the southern and western
Gulf. Fish and algae aquaculture, on the other hand, remains marginal.
Considering this, we only included invertebrates aquaculture for the
aquaculture driver layer. However, if fish or algae farming were to become more
important, these drivers should be incorporated in future analyses as individual
layers, as impacts vary between types of aquaculture.

Aquaculture activities are highly localized and potential effects do not or
rarely extend beyond the location of the farms. We therefore only considered
the location of sites to characterize the distribution of this driver.
We were unable to characterize site production in terms of biomass farmed, which
could provide an indication of the intensity of aquaculture activities. As such,
we considered aquaculture as binary presence-absence data.

@@

Terrestrial stable lights at night mostly represent light from human settlements and industrial sites with electricity. We thus used lights at night as a proxy of coastal infrastructure development.

@@

As in @halpern2008a and @halpern2015, we used the sum of coastal populations as
a proxy of direct human impact. We used Statistics Canada dissemination area
population count from the 2016 census to obtain coastal population size
around the St. Lawrence [@statistics-canada2017]. Dissemination areas are the smallest
standard geographic area in which census data are disseminated. They combine
to cover all of Canada and are highly variable in shapes and sizes.
\textcolor{blue}{For example, their sizes range from 0.0003 to over 44000 $km^2$ in our study area.}
The census provides population count within the boundary
of each dissemination area, which we used to evaluate total coastal population.

@@

To characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multiplied the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

@@

# ~~Threat complexes~~ \textcolor{blue}{Cumulative exposure profiles}
